\chapter{State of the Art}
\label{chap:state_of_the_art}

\section*{Introduction}
\addcontentsline{toc}{section}{Introduction}
Before trying to make our own solution, we must understand whats already been done in both the clinical and a.i (deep reinforcement learning mostly) fields, and technical background that supports the
work. The basic idea is simple: if we want to build a useful endoscope agent, we have
to understand the medical workflow, the \textbf{vision methods} that read the scene, and the
\textbf{learning methods} that decide what to do next.

In this context, clinical guidance on \textbf{colorectal polypectomy} and endoscopic resection
gives us the safety baseline \cite{ferlitsch2017esge,kaltenbach2020usmstf}. At the same
time, recent clinical trials show that real-time computer-aided detection can improve
lesion detection during routine colonoscopy \cite{wang2019cade,repici2020cade}.

The chapter moves from the clinical context to the learning methods used in intelligent
systems. Along the way, we keep an eye on the practical things that matter most here:
\textbf{data quality}, \textbf{safety}, \textbf{robustness}, and \textbf{evaluation}. Those are not just side details; they
shape the design choices in the rest of the document.

\textbf{Chapter roadmap:} We start with colorectal polyps and polypectomy, then move to
Artificial Intelligence as the broader field. After that, we cover deep reinforcement learning
as the main framework for sequential decision-making, and we review the neural
architectures used for perception and temporal modeling (\textbf{CNN} and \textbf{LSTM}). We end with
Proximal Policy Optimization (\textbf{PPO}), the policy-gradient method used in the simulation.
\input{chapters/101_polypectomy}
\input{chapters/102_artificial_intelligence}
\input{chapters/103_RL}
\input{chapters/104_CNN}
\input{chapters/105_LSTM}
\input{chapters/106_PPO}

\section*{Conclusion}
\addcontentsline{toc}{section}{Conclusion}
The state of the art makes one thing pretty clear: reliable assistance in endoscopy
needs both clear \textbf{perception} and reliable \textbf{decision-making}. By
perception we mean the ability of vision models to correctly see and describe the
scene (for example: detect polyps, identify blood or debris, and recognise tools).
These models must cope with variation between patients, different endoscope devices
and settings, changing lighting, uneven bowel preparation, occlusions (fluids or
debris), and motion blur from camera movement. By decision-making we mean the
policies that select actions over time; these policies must remain sensible when
observations are noisy, avoid risky moves, and behave in ways clinicians expect.

The literature highlights several clear challenges. First, rare but important events
are difficult to collect in large numbers, so training data can be limited or biased.
Second, producing accurate annotations is time-consuming and sometimes inconsistent
between annotators. Third, models trained in one hospital or on one device often do
not generalise well to other hospitals or devices. Finally, common metrics like
accuracy do not capture clinical risk, so evaluation must include safety-focused
measures (for example: missed-lesion rate, number of unsafe actions, and adherence
to clinical guidelines). To address these gaps, this work focuses on controlled
training environments, careful reward design that penalises unsafe behaviour, and
task-level metrics that measure both \textbf{performance} (detection and task success)
and \textbf{safety} (risk of harm, false negatives, and guideline compliance).
