\section{Proximal Policy Optimization (PPO)}
\section{Proximal Policy Optimization (PPO)}
\label{sec:ppo}

\subsection*{Introduction}
Proximal Policy Optimization (PPO) is a state-of-the-art policy gradient algorithm that balances sample efficiency, implementation simplicity, and training stability. Introduced by OpenAI in 2017, PPO has become one of the most widely used algorithms in modern reinforcement learning due to its empirical performance across diverse domains and its robustness to hyperparameter choices.

The canonical PPO formulation and clipped objective are described by Schulman et al. \cite{schulman2017ppo}.

This chapter explains why PPO became so popular, what actually makes it work, and how it compares with earlier policy gradient methods. It helps to already know the basics of policy gradients from the previous chapter.

PPO is especially popular in simulation-based robotics toolchains because it gives a practical trade-off: stable updates without the heavy implementation cost of strict trust-region solvers. That makes it a good fit for iterative experiments where the environment, observation space, and reward design keep changing.

\subsection{Motivation: Trust Region Policy Optimization}
Early policy gradient methods suffered from instability due to large and destructive policy updates. If the policy changes too drastically in a single update, the value estimates become stale and learning diverges. Trust Region Policy Optimization (TRPO) addressed this by constraining policy updates to remain within a small trust region \cite{schulman2015trpo}:

$$\text{maximize} \quad \mathbb{E}\left[\frac{\pi_\theta(a|s)}{\pi_{\theta_\text{old}}(a|s)} \hat{A}_t\right]$$
$$\text{subject to} \quad \mathbb{E}\left[\text{KL}\left(\pi_{\theta_\text{old}}(\cdot|s), \pi_\theta(\cdot|s)\right)\right] \leq \delta$$

where $\hat{A}_t$ is the advantage estimate and $\delta$ is the trust region radius.

While effective, TRPO is complex to implement and computationally expensive. PPO simplifies this approach while retaining the key benefit of controlled policy updates.

\subsection{The PPO-Clip Objective}
PPO replaces the hard KL divergence constraint with a soft penalty implemented via a clipping mechanism. The clipped objective is:

$$L^{\text{CLIP}}(\theta) = \mathbb{E}\left[\min\left(r_t(\theta) \hat{A}_t, \text{clip}(r_t(\theta), 1-\epsilon, 1+\epsilon) \hat{A}_t\right)\right]$$

where $r_t(\theta) = \frac{\pi_\theta(a_t|s_t)}{\pi_{\theta_\text{old}}(a_t|s_t)}$ is the probability ratio, and $\epsilon$ (typically $0.1$ to $0.2$) is the clipping range.

The clipping mechanism prevents the probability ratio from moving too far from $1$:

\begin{itemize}

\item If the advantage is positive and the ratio exceeds $1 + \epsilon$, the objective becomes constant, removing incentive for further increase.

\item If the advantage is negative and the ratio falls below $1 - \epsilon$, the objective plateaus, preventing excessive reduction.

\item If the ratio stays within the range $[1-\epsilon, 1+\epsilon]$, the update is unclipped and proceeds normally.

\end{itemize}

This design automatically adapts the step size: large advantages tolerate larger steps, while small advantages keep updates conservative.

\subsection{Actor-Critic Architecture}
PPO uses an actor-critic architecture combining a policy (actor) and a value function (critic):

\begin{itemize}

\item \textbf{Actor:} Parametrized by $\theta$, outputs the policy $\pi_\theta(a|s)$. Updated using the clipped policy gradient objective.

\item \textbf{Critic:} Parametrized by $\phi$, outputs the value estimate $V_\phi(s)$. Updated to minimize the mean squared error between predicted and target returns.

\end{itemize}

The critic provides advantage estimates:

$$\hat{A}_t = r_t + \gamma V_\phi(s_{t+1}) - V_\phi(s_t)$$

A good value function estimate is crucial for reducing the variance of policy gradient estimates while maintaining low bias.

In practice, many PPO implementations share a common network trunk with separate policy and value heads, allowing efficient use of learned representations.

\subsection{Algorithm and Implementation Details}
PPO operates on batches of trajectory data collected by running the current policy in the environment. The typical training loop is:

\begin{enumerate}

\item \textbf{Collect Experience:} Run the policy for $N$ timesteps (or $M$ episodes), recording states, actions, rewards, and log probabilities.

\item \textbf{Compute Advantages:} Use the critic to estimate advantages, typically with Generalized Advantage Estimation (GAE) for variance reduction \cite{schulman2016gae}:
$$\hat{A}_t = \sum_{l=0}^{T-t-1} (\gamma \lambda)^l \left(r_{t+l} + \gamma V_\phi(s_{t+l+1}) - V_\phi(s_{t+l})\right)$$
where $\lambda \in [0,1]$ controls the bias-variance tradeoff.

\item \textbf{Update Policy and Value Function:} Perform multiple gradient update passes (epochs) over the collected data, updating both the actor and critic. Data is typically split into minibatches.

\item \textbf{Repeat:} Collect new data and repeat until convergence.

\end{enumerate}

Key implementation considerations:

\begin{itemize}

\item PPO typically uses first-order optimization (e.g., Adam) rather than second-order methods \cite{kingma2014adam}.

\item Value function loss can be combined with policy loss: $L(\theta,\phi) = L^{\text{CLIP}}(\theta) + c_1 L^{\text{VF}}(\phi) + c_2 H(\pi_\theta)$, where $H$ is entropy regularization.

\item Learning rates are often scheduled to decay over training.

\item Gradient clipping prevents exploding gradients.

\begin{table}[h]
\centering
\caption{Key PPO hyperparameters and typical values}
\begin{tabular}{|l|l|l|}
\hline
	extbf{Hyperparameter} & \textbf{Range} & \textbf{Effect} \\ \hline
Clipping range ($\epsilon$) & 0.1--0.3 & Controls step size \\ \hline
Discount factor ($\gamma$) & 0.95--0.99 & Future reward weight \\ \hline
GAE $\lambda$ & 0.9--0.99 & Bias-variance tradeoff \\ \hline
Learning rate & $10^{-3}$ to $10^{-5}$ & Optimization step size \\ \hline
Batch size & 64--4096 & Gradient quality vs memory \\ \hline
Epochs per update & 3--10 & Gradient update passes \\ \hline
Entropy coeff. & 0--0.01 & Exploration bonus \\ \hline
\end{tabular}
\label{tab:ppo_hyperparams}
\end{table}

	extit{[Placeholder for Figure: PPO training algorithm flowchart]}

\subsection{Advantages Over Prior Methods}
PPO offers several advantages that have made it a common choice for many applications:

\begin{itemize}

\item \textbf{Simplicity:} The clipped objective is easy to implement and understand, requiring only a few lines of code.

\item \textbf{Sample Efficiency:} By reusing data across multiple gradient steps, PPO requires fewer environment interactions than on-policy methods like REINFORCE.

\item \textbf{Stability:} The clipping mechanism and adaptive learning provide robust training across diverse tasks and hyperparameters.

\item \textbf{Scalability:} PPO scales well to high-dimensional state and action spaces and has been successfully applied to continuous control and large language models.

\item \textbf{Robustness:} PPO is less sensitive to hyperparameter tuning than TRPO or earlier policy gradients.

\end{itemize}

\subsection{Applications and Performance}
PPO has achieved state-of-the-art results across numerous domains:

\begin{itemize}

\item \textbf{Robotics:} Continuous control tasks such as locomotion, manipulation, and navigation.

\item \textbf{Games:} Video games (Atari, Dota 2) and simulated environments.

\item \textbf{Large Language Models:} RLHF (Reinforcement Learning from Human Feedback) training for language model fine-tuning uses PPO variants.

\item \textbf{Autonomous Systems:} Vehicle control and navigation tasks.

\end{itemize}

	extit{[Placeholder for Table: Performance Comparison of PPO against TRPO, A3C, DDPG, and other RL algorithms]}

	extit{[Placeholder for Figure: Example PPO learning curves on standard benchmarks (e.g., MuJoCo continuous control environments)]}

\begin{table}[h]
\centering
\caption{PPO vs. competing policy gradient algorithms}
\begin{tabular}{|l|l|l|l|}
\hline
	extbf{Algorithm} & \textbf{Sample Efficiency} & \textbf{Stability} & \textbf{Ease of Use} \\ \hline
PPO & High & Very High & Very Easy \\ \hline
TRPO & High & High & Difficult \\ \hline
A3C & Medium & Medium & Moderate \\ \hline
DDPG & Medium & Low & Moderate \\ \hline
SAC & High & High & Moderate \\ \hline
REINFORCE & Low & Low & Easy \\ \hline
\end{tabular}
\label{tab:ppo_comparison}
\end{table}

\begin{table}[h]
\centering
\caption{PPO performance on benchmark environments}
\begin{tabular}{|l|c|l|}
\hline
	extbf{Benchmark} & \textbf{Final Performance} & \textbf{Convergence Time} \\ \hline
MuJoCo Walker2d & $\sim$5000 & 1--2M steps \\ \hline
MuJoCo HalfCheetah & $\sim$5000 & 1--2M steps \\ \hline
Atari (avg.) & 1500\% baseline & 2--10M steps \\ \hline
Robotics (success) & $>$95\% & 5--20M steps \\ \hline
\end{tabular}
\label{tab:ppo_benchmarks}
\end{table}

	extit{[Placeholder for Figure: PPO learning curves on MuJoCo continuous control tasks]}

	extit{[Placeholder for Figure: Visual examples of PPO applications (robot locomotion, game playing, etc.)]}

\subsection{Limitations and Open Questions}
Despite its success, PPO has limitations:

\begin{itemize}

\item \textbf{Off-Policy Learning:} PPO is on-policy and discards old data. Extensions for off-policy learning have been developed but remain less stable.

\item \textbf{Value Function Bootstrapping:} Reliance on value function estimates can propagate errors, especially early in training.

\item \textbf{Hyperparameter Sensitivity:} While more robust than some methods, performance can still depend significantly on choices like clipping range $\epsilon$ and discount factor $\gamma$.

\item \textbf{Exploration:} PPO relies on entropy regularization for exploration; more sophisticated exploration strategies are an active research area.

\end{itemize}

\subsection{Summary}
Proximal Policy Optimization is one of the reasons policy-gradient methods became much more practical. It mixes the stability ideas from trust-region methods with a simple clipped objective, so it works well in many simulation-based control problems.

Here, PPO is useful because it gives a stable baseline for learning sequential decision policies in controlled experiments. When it is paired with perception modules like CNN-based visual encoders and with careful reward design, it supports iterative development and evaluation before any more serious deployment step.
\end{flushleft}
