\chapter{General Conclusion}
\label{chap:general_conclusion}

\section*{Conclusion}
\begin{flushleft}
This thesis explored the absolutely fascinating intersection of \textbf{computer vision}, \textbf{deep reinforcement learning}, and \textbf{endoscopic surgery} to tackle one of the trickiest procedures in modern medicine: \textbf{colorectal polypectomy}. Did you know the human colon has an internal surface area of roughly 2,000\,cm\textsuperscript{2}? Navigating that kind of real estate and carefully getting a snare around a polyp is far beyond a simple image classification task. It's actually a wildly complex, step-by-step decision problem because every single move completely changes what the camera sees next!

To solve this, we built a highly sophisticated, physics-driven \textbf{Unity simulation} to train vision-only agents. We split the procedure into two separate parts: \textbf{endoscope navigation} and \textbf{snare capture}. This simple trick took a huge weight off the AI's shoulders. And the results speak for themselves! In the first phase, our agent learned to safely steer the endoscope tip through those tortuous haustral folds, actively dodging collisions by reading real-time visual cues like \textbf{specular highlights} and dynamic shadows, achieving a rock-solid \textbf{best score of 14.67} after 6.5 million steps. In the second phase, the agent mastered the incredibly delicate art of looping a snare around a pedunculated polyp. This was all thanks to a carefully engineered \textbf{``breadcrumb'' reward system} and an \textbf{Intrinsic Curiosity Module (ICM)}. Both phases relied heavily on \textbf{Proximal Policy Optimization (PPO)} paired with \textbf{Convolutional Neural Networks (CNNs)} and \textbf{LSTM memory} to deliver remarkably safe and reliable policies.

Ultimately, the core contribution of this work is proving that we can build a software-first, structured foundation for \textbf{adaptive, semi-autonomous systems} that can assist doctors in totally chaotic operating environments. 
\end{flushleft}

\section{Limitations}

As much as We are content with this project, let's be totally honest about the hurdles. First off, our physics are heavily simplified. We haven't implemented true \textbf{soft-tissue deformation}, which means the colon walls don't squish and stretch realistically when the endoscope bumps into them. Secondly, under the hood, our polyps are basically clever \textbf{geometric sphere-proxies} rather than fully anatomical soft-body simulations. Finally, training a brilliant policy in a simulator is one thing, but throwing it into a real, messy clinical setting is a whole different beast. AI researchers call this the \textbf{sim-to-real gap}, and it's a massive hurdle that regularly breaks even the toughest Deep RL models.

\section{Future Work}

Looking ahead, there are some incredibly exciting avenues to explore! For future work, it would be absolutely amazing to integrate a high-end biomechanical engine like \textbf{SOFA (Simulation Open Framework Architecture)} to handle hyper-realistic tissue deformation. Another super cool idea would be to bolt on a dedicated \textbf{object-detection AI} (like YOLO) to spot the polyps automatically. This would create a powerful \textbf{hybrid system} where the computer vision model tells the Deep RL agent exactly where to aim! Bridging that reality gap with richer clinical datasets and eventually testing these intelligent policies on real-world \textbf{robotic endoscopy hardware} is the ultimate dream.
